Reasoning About a Machine with Local Capabilities

### 2018-05-09 toplas/paper.tex
Added a figure that illustrates the last example for when the adv hides a stack pointer on the stack.

      % Stack pointer
      \begin{scope}
        \clip (4.6,5) rectangle (7.6,11.5);
        \capbrace[sp]{(4.6,5)}{(4.6,12)}
      \end{scope}
      \draw node[right=0cm of sp] { sp };
    \end{tikzpicture}
    \caption{}
  \label{fig:illu-intro}
  \end{subfigure}
  \begin{subfigure}{0.333\linewidth}
    \begin{tikzpicture}[scale=.5, every node={scale=.5}]
    % recurrent parts
    \stdstackstart

@@

      % Stack pointer
      \begin{scope}
        \clip (4.6,2) rectangle (7.6,11.5);
        \capbrace[sp]{(4.6,2)}{(4.6,12)}
      \end{scope}
      \draw node[right=0cm of sp] { sp };
    \end{tikzpicture}
  \caption{}
  \label{fig:illu-priv-data}
  \end{subfigure}
  \begin{subfigure}{0.333\linewidth}
    \begin{tikzpicture}[scale=.5, every node={scale=.5}]
    % recurrent parts
    \stdstackstart

@@

      % Stack pointer
      \begin{scope}
        \clip (4.6,2) rectangle (7.6,11.5);
        \capbrace[sp]{(4.6,2)}{(4.6,12)}
      \end{scope}
      \draw node[right=0cm of sp] { sp };
    \end{tikzpicture}
  \caption{}
  \label{fig:illu-clear-priv}
  \end{subfigure}

@@

      % Stack pointer
      \begin{scope}
        \clip (4.6,5) rectangle (7.6,11.5);
        \capbrace[sp]{(4.6,5)}{(4.6,13)}
      \end{scope}
      \draw node[right=0cm of sp] { sp };
    \end{tikzpicture}
  \caption{}
  \label{fig:illu-call-adv}
  \end{subfigure}
  \begin{subfigure}{0.333\linewidth}
    \begin{tikzpicture}[scale=.5, every node={scale=.5}]
    % recurrent parts
    \stdstackstart

@@

      % Stack pointer
      \begin{scope}
        \clip (4.6,2) rectangle (7.6,11.5);
        \capbrace[sp]{(4.6,2)}{(4.6,13)}
      \end{scope}
      \draw node[right=0cm of sp] { sp };
    \end{tikzpicture}
  \caption{}
  \label{fig:illu-adv-data}
  \end{subfigure}
  \begin{subfigure}{0.333\linewidth}
    \begin{tikzpicture}[scale=.5, every node={scale=.5}]
    % recurrent parts
    \stdstackstart

@@

      % Stack pointer
      \begin{scope}
        \clip (4.6,2) rectangle (7.6,11.5);
        \capbrace[sp]{(4.6,2)}{(4.6,13)}
      \end{scope}
      \draw node[right=0cm of sp] { sp };
    \end{tikzpicture}
  \caption{}
  \label{fig:illu-clear-all}
  \end{subfigure}
  \caption{Illustration of the situations related to stack clearing when returning to an adversary.
    The greyed out areas are parts of the stack that are not accessible at the time. }
  \label{fig:ret-adv}
\end{figure}
\textbf{Invoking the adversary} A slightly more complex case is invoking the
adversary. As above, we clear all the non-argument registers, as well as the
part of the stack that we are not using (because, as above, it may contain local
capabilities from previously executed code that the adversary could exploit in
the same way). We leave a copy of the stack pointer in $r_\stk$, but only after
we have used the $\subseg{}{}{}$ instruction to shrink its authority to the part
that we are not using ourselves.

@@

Note that storing stack pointers for use during future invocations would also be
dangerous in itself, i.e. not just because it can be used to store return
pointers. Imagine the adversary stores their stack pointer, invokes trusted code
that uses part of the stack to store private data and then invokes the adversary
again with a stack pointer restricted to exclude the part containing the private
data. If the adversary had a way of keeping hold of their old stack pointer, it
could access the private data stored there by the trusted code and break
local-state encapsulation.  
% \lau{Do we say anything about arguments to the callback? Specifically, it should not be okay to uncritically pass on argument we recieved from an adversary. Can we even to any extend allow local arguments that we pass on from an unknown source?}
% Domi: this is now discussed in the discussion section, no need to discuss it here?

>>>>

      % Stack pointer
      \begin{scope}
        \clip (4.6,5) rectangle (7.6,11.5);
        \capbrace{(4.6,5)}{(4.6,12)}{sp}
      \end{scope}
    \end{tikzpicture}
    \caption{}
  \label{fig:illu-intro}
  \end{subfigure}
  \begin{subfigure}{0.333\linewidth}
    \begin{tikzpicture}[scale=.5, every node={scale=.5}]
    % recurrent parts
    \stdstackstart

@@

      % Stack pointer
      \begin{scope}
        \clip (4.6,2) rectangle (7.6,11.5);
        \capbrace{(4.6,2)}{(4.6,12)}{sp}
      \end{scope}
    \end{tikzpicture}
  \caption{}
  \label{fig:illu-priv-data}
  \end{subfigure}
  \begin{subfigure}{0.333\linewidth}
    \begin{tikzpicture}[scale=.5, every node={scale=.5}]
    % recurrent parts
    \stdstackstart

@@

      % Stack pointer
      \begin{scope}
        \clip (4.6,2) rectangle (7.6,11.5);
        \capbrace{(4.6,2)}{(4.6,12)}{sp}
      \end{scope}
    \end{tikzpicture}
  \caption{}
  \label{fig:illu-clear-priv}
  \end{subfigure}

@@

      % Stack pointer
      \begin{scope}
        \clip (4.6,5) rectangle (7.6,11.5);
        \capbrace{(4.6,5)}{(4.6,13)}{sp}
      \end{scope}
    \end{tikzpicture}
  \caption{}
  \label{fig:illu-call-adv}
  \end{subfigure}
  \begin{subfigure}{0.333\linewidth}
    \begin{tikzpicture}[scale=.5, every node={scale=.5}]
    % recurrent parts
    \stdstackstart

@@

      % Stack pointer
      \begin{scope}
        \clip (4.6,2) rectangle (7.6,11.5);
        \capbrace{(4.6,2)}{(4.6,13)}{sp}
      \end{scope}
    \end{tikzpicture}
  \caption{}
  \label{fig:illu-adv-data}
  \end{subfigure}
  \begin{subfigure}{0.333\linewidth}
    \begin{tikzpicture}[scale=.5, every node={scale=.5}]
    % recurrent parts
    \stdstackstart

@@

      % Stack pointer
      \begin{scope}
        \clip (4.6,2) rectangle (7.6,11.5);
        \capbrace{(4.6,2)}{(4.6,13)}{sp}
      \end{scope}
    \end{tikzpicture}
  \caption{}
  \label{fig:illu-clear-all}
  \end{subfigure}
  \caption{Illustration of the situations related to stack clearing when returning to an adversary.
    The greyed out areas are parts of the stack that are not accessible at the time.}
  \label{fig:ret-adv}
\end{figure}
\textbf{Invoking the adversary} A slightly more complex case is invoking the
adversary. As above, we clear all the non-argument registers, as well as the
part of the stack that we are not using (because, as above, it may contain local
capabilities from previously executed code that the adversary could exploit in
the same way). We leave a copy of the stack pointer in $r_\stk$, but only after
we have used the $\subseg{}{}{}$ instruction to shrink its authority to the part
that we are not using ourselves.

@@

Note that storing stack pointers for use during future invocations would also be
dangerous in itself, i.e. not just because it can be used to store return
pointers. Imagine the adversary stores their stack pointer (\ref{fig:illu-adv-store-sp}), invokes trusted code
that uses part of the stack to store private data (\ref{fig:illu-adv-call}) and then invokes the adversary
again with a stack pointer restricted to exclude the part containing the private
data (\ref{fig:illu-adv-called}). If the adversary had a way of keeping hold of their old stack pointer, it
could access the private data stored there by the trusted code and break
local-state encapsulation.  
% \lau{Do we say anything about arguments to the callback? Specifically, it should not be okay to uncritically pass on argument we recieved from an adversary. Can we even to any extend allow local arguments that we pass on from an unknown source?}
% Domi: this is now discussed in the discussion section, no need to discuss it here?

    % attacker ..
      \actadv{(0,2)}{(4.5,5)}{Adv. stack frame}
      \draw[fill=white] (0,10) rectangle (4.5,10.5) node[pos=.5,color=red] {\footnotesize sp};
    % Stack pointer
      \begin{scope}
        \clip (4.6,2) rectangle (7.6,11.5);
        \capbracebot{(4.6,2)}{(4.6,12)}{sp}
      \end{scope}
    \end{tikzpicture}
    \caption{}
  \label{fig:illu-adv-store-sp}
  \end{subfigure}
  \begin{subfigure}{0.333\linewidth}
    \begin{tikzpicture}[scale=.5, every node={scale=.5}]
    % recurrent parts
    \stdstackstart

    % attacker ..
      \inactadv{(0,2)}{(4.5,5)}{Adv. stack frame}
      \draw[fill=white] (0,10) rectangle (4.5,10.5) node[pos=.5,color=red] {\footnotesize sp};

    % Our stack frame
      \actsf{(0,5)}{(4.5,6.5)}{Our stack frame}

    % Stack pointer
      \begin{scope}
        \clip (4.6,2) rectangle (7.6,11.5);
        \capbracebotred{(4.6,2)}{(4.6,12)}{sp}
      \end{scope}
      \begin{scope}
        \clip (5.6,5) rectangle (9,11.5);
        \capbrace{(5.6,5)}{(5.6,12)}{sp'}
      \end{scope}

    \end{tikzpicture}
  \caption{}
  \label{fig:illu-adv-call}
  \end{subfigure}
  \begin{subfigure}{0.333\linewidth}
    \begin{tikzpicture}[scale=.5, every node={scale=.5}]
    % recurrent parts
    \stdstackstart

    % Stored stack pointer
      \draw[fill=white] (0,10) rectangle (4.5,10.5) node[pos=.5,color=red] {\footnotesize sp};

    % Our stack frame
      \inactsf{(0,5)}{(4.5,6.5)}{Our stack frame}

    % Second attacker
      \actadv{(0,6.5)}{(4.5,9.5)}{Callee \\ \footnotesize stack frame}

    % Stack pointer
      \begin{scope}
        \clip (4.6,2) rectangle (7.6,11.5);
        \capbracebotred{(4.6,2)}{(4.6,12)}{sp}
      \end{scope}
      \begin{scope}
        \clip (5.6,6.6) rectangle (9,11.5);
        \capbrace{(5.6,6.5)}{(5.6,12)}{sp'}
      \end{scope}
    \end{tikzpicture}
  \caption{}
  \label{fig:illu-adv-called}
  \end{subfigure}
  \caption{Illustration of the situations related to stack clearing when invoking an adversary.
    The greyed out areas are parts of the stack that are not accessible at the time.
  }
  \label{fig:ret-adv}
\end{figure}

### 2018-05-09 toplas/preamble-illu.tex
Added a figure that illustrates the last example for when the adv hides a stack pointer on the stack.

\newcommand{\capbrace}[3][sp1]{
  \draw [decorate,decoration={brace,amplitude=10pt,mirror,raise=4pt},yshift=0pt]
  #2 -- #3 node[draw=black] (#1) [black,midway,xshift=0.8cm] {};

}

>>>>

\newcommand{\capbrace}[4][sp1]{
  \draw [decorate,decoration={brace,amplitude=10pt,mirror,raise=4pt},yshift=0pt]
  #2 -- #3 node (#1) [black,midway,xshift=0.8cm] {\footnotesize #4};
}

\newcommand{\capbracebot}[4][sp1]{
  \draw [decorate,decoration={brace,amplitude=10pt,mirror,raise=4pt,aspect=0.25},yshift=0pt]
  #2 -- #3 node (#1) [black,near start,xshift=0.8cm] {\footnotesize #4};
}

\newcommand{\capbracebotred}[4][sp1]{
  \draw [decorate,red,decoration={brace,amplitude=10pt,mirror,raise=4pt,aspect=0.25},yshift=0pt]
  #2 -- #3 node (#1) [red,near start,xshift=0.8cm] {\footnotesize #4};
}
